Sensitivity to measurement error in dissolved gas measurements

### 2024-05-24 scripts/DG_sensitivity_to_measurement_error.Rmd
Updated formulas and figures; Added scenarios

---
title: "Evaluating sensitivity to measurement error in dissolved gas measurements"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
  html_notebook:
    depth: 4
    fig_caption: true
    number_sections: true
    toc: true
    toc_float: true
link-citations: true
---

<style>
.vscroll-plot {
    width: 800px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

# Background
In this document we explore a range of models for fitting dissolved $\text{N}_2\text{O}$ data observed as part of the 2017 National Lakes Assessment (NLA).

# Setup R
```{r setup_1, include=FALSE, echo=FALSE}
library(knitr)
# opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
# 
# #set directories and R package library
# setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
# .libPaths("C:/Users/rmartin/Desktop/R/library")
```

```{r load_packages, warning=FALSE, message=FALSE}
library(brms)
library(bayesplot)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(readxl)
library(janitor)
library(sf)

options(mc.cores = parallel::detectCores())
options( max.print = 2000 )
```

```{r utility_functions, include=FALSE}
# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

@@

#Calc geometric mean
gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(mean(log(x), na.rm = na.rm))
  } else {
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

gm_sd = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(sd(log(x), na.rm = na.rm))
  } else {
    exp(sd(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

#calculate skew
MySkew <- function(x) {
     n <- length(x)
     v <- var(x)
     m <- mean(x)
     third.moment <- (1/(n - 2)) * sum((x - m)^3)
     third.moment/(var(x)^(3/2))
 }

#Center and scale by (1) sd
MyStd <- function(x){ (x-mean(x))/(sd(x)*2)}

#First letter in string to uppercase
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x}
```

# Dissolved gas concentration calculation

The full equation for estimating the concentration (nM) of $N_20$ gas dissolved in water is:

$$C = 1^{-6}B \frac{V_{air}(Mr_h - Mr_r)}{(8.3144598 * K * V_{water}) + H^{\theta} \exp( 2700 * ( \frac{1}{K} - \frac{1}{298.15}) * Mr_h}$$

# Set up model for simulating true value

```{r setvars_and_check}
load(paste0(localPath, "\\Environmental Protection Agency (EPA)",
            "\\ORD NLA17 Dissolved Gas - Documents\\",
            "inputData\\dg.2021-02-01.RData"))

# summary of measured N2O ppm in equilibrated headspace gas
dg %>%
  group_by(sample.source) %>%
  summarize(mean_n2O = mean(n2o.ppm.gc, na.rm = TRUE),
          median_n2o = median(n2o.ppm.gc, na.rm = TRUE)) #mean = 0.36, median = 0.31

#    mean_n2o  median_n2o
#AIR	0.3093312	0.3104941		
#DG	 0.3618681	0.3067609

B <- 99 # barometric pressure (kPa)
V_g <- 35 # volume of reference gas in headspace
V_w <- 105 # volume of water below headspace
T_c <- 23 # temperature ( C )
C_e <- 0.307 # gas mixture ratio in headspace
C_a <- 0.310 # gas mixture ratio in reference
H <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant

C <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )

@@

round( C, 1 )

```
## Define GC measurement error
Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.

```{r get_gc_error}
# READ GC DATA----------------
##########################################################################
#       Code is commented out because collaborators may not have access  #
#       to the file paths. The calculated standard deviation is hard     #
#       coded below.
##########################################################################

@@

## Simulate measurement error in GC measurement of N2O in equilibrated headspace
```{r simulate_me_Ce, message=F}
C_me <- rep( NA, 1e4 ) # "me" is "measured" gas in headspace. empty vector

# this simulates measurement error on GC.
# assume TRUE value is C_e and error is simulated with with sd of replicate GC measurements
C_e_me <- rnorm( 1e4, C_e, n2o_sd ) # me in measured gas in headspace.
                                   # add random error to measured concentration

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, C_e_me = C_e_me)

ggplot(C_df, aes(x = C_e_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_e , color = 'red' ) + 
  ggtitle( "Simulated measurement error in equilibrated gas concentration") +
  xlab( "concentration (nmol L-1)")

ggplot(C_df, aes(x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof equilibrated headspace") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes(x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof equilibrated headspace") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error for GC measurement of N2O in air
```{r simulate_me_Ca, message=F}
C_me <- rep( NA, 1e4 )

C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
Ca_df <- tibble(C_me = C_me, C_a_me = C_a_me)

ggplot(Ca_df, aes(x = C_a_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (?)")

ggplot(Ca_df, aes(x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof N2O in air") +
  xlab( "concentration (nMol)")

ggplot(Ca_df, aes(x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N2O concentration due to GC measurment error \nof N2O in air") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in $C_e$ and $C_a$
```{r simulate_me_Ca_Ce, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me)

ggplot(C_df, aes(x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof 
           N2O in equilibrated headspace and air") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes(x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof N2O in equilibrated headspace and air") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in water temperature
```{r simulate_me_T_c, message=F}
C_me <- rep( NA, 1e4 )

# How to get standard deviation of temperature measurement?
# calibration certificate reports an uncertainty of +/-0.058C
# take standard deviation of simulated measurements with specified uncertainty
T_c_sd <- sd(runif(n = 1e4, min = T_c - 0.058, max = T_c + 0.058)) # calibration certificate reports an uncertainty of +/-0.058C
T_c_me <- rnorm( 1e4, T_c, T_c_sd ) #m.e. in measured water temp

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, T_c_me = T_c_me)

ggplot(C_df, aes(x = T_c_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = T_c , color = 'red' ) + 
  ggtitle( "Simulated measurement error in temperature") +
  xlab( "temperature (C)")

ggplot(C_df, aes( x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to water temperature \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to water temperature \nmeasurment error") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in barometric pressure
```{r simulate_me_B, message=F}
C_me <- rep( NA, 1e4 )

# How to get standard deviation of barometric pressure?
# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa)
# take standard deviation of simulated measurements with specified uncertainty
B_sd <- sd(runif(n = 1e4, min = B - 0.4, max = B + 0.4)) # calibration certificate reports an uncertainty of +/-0.058C
B_me <- rnorm( 1e4, B, 0.25 ) #m.e. in measured barometric pressure

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, B_me = B_me, B_ru = runif(n = 1e4, min = B - 0.4, max = B + 0.4))

ggplot(C_df, aes( x = B_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = B , color = 'red' ) + 
  ggtitle( "Simulated measurement error in barometric pressure") +
  xlab( "barometric pressure (kPa)")

ggplot(C_df, aes( x = B_ru )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = B , color = 'red' ) + 
  ggtitle( "Simulated measurement error in barometric pressure") +
  xlab( "barometric pressure (kPa)")

ggplot(C_df, aes( x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to barometric pressure \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to barometric pressure \nmeasurment error") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in water volume
```{r simulate_me_V_w, message=F}
C_me <- rep( NA, 1e4 )

V_w_me <- rnorm( 1e4, V_w, 1 ) #m.e. in measured water volume
V_g_me <- 140 - V_w_me

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g_me[ i ] * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, V_w_me = V_w_me)

ggplot(C_df, aes( x = V_w_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = V_w , color = 'red' ) + 
  ggtitle( "Simulated measurement error in water volume") +
  xlab( "water vol (mL)")

ggplot(C_df, aes( x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissollved N2O concentration due to water volume \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N2O concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error for all observables
```{r simulate_me_all, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me)

ggplot(C_df, aes( x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors") +
  xlab( "absolute error (nMol)")

>>>>

---
title: "Sensitivity to measurement error in dissolved gas measurements"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 5
    number_sections: true
    keep_html: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, warning=FALSE, message=FALSE}
library(knitr)
library(rmarkdown)
library(bayesplot)
library(kableExtra)
library(ggpubr)
library(tidyverse)
library(tidybayes)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(readxl)
library(janitor)
library(sf)

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

@@

options(max.print = 2000)
```

# Background

In this document we explore a range of scenarios wherein measurement errors 
could arise in methods for determining the concentration of $\text{N}_2\text{O}$ 
from field and laboratory measurements observed as part of the 2017 National 
Lakes Assessment (NLA).

# Dissolved concentration

The equation for estimating the concentration (nM) of $\text{N}_2\text{O}$ gas
dissolved in water is:

$$C = B^{-6} \cdot \left(\frac{V_g(M_h - M_r)}{(8.3144598 \cdot K \cdot V_w) + (H^{\theta} \cdot e^{2700 \cdot ( \frac{1}{K} - \frac{1}{298.15})} \cdot M_h)} \right)$$

where:

-   $C$ = concentration (nM)
-   $B$ = barometric pressure (kPa)
-   $V_g$ = volume of reference gas in headspace (mL)
-   $M_h$ = gas mixture ratio in the headspace (C_e below)
-   $M_r$ = gas mixture ratio in the reference (C_a_n2o below)
-   $V_w$ = volume of water below the headspace (mL)
-   $K$ = temperature (K) (temp celsius or T_c + 273 below)
-   $H^{\theta}$ = Henry's law constant ($H_{N_2O}$ or $H_{Ar}$ below)

and:

$$\begin{split}
 V & = V_g + V_w \\
  V_g & = V \cdot R_M \\
  V_w & = V \cdot (1 - R_M)
  \end{split}$$

where:

-   $V$ = Total volume of the mixing vessel (e.g. 140mL syringe)
-   $V_g$ = Volume of gas (mL) in the mixing vessel 
-   $V_w$ = Volume of water (mL) in the mixing vessel
-   $R_M$ = Mixing ratio (gas:water)

## Simulate $\text{N}_2\text{O}$

```{r summary_empirical, echo=TRUE}
load(paste0(localPath, "\\Environmental Protection Agency (EPA)",
            "\\ORD NLA17 Dissolved Gas - Documents\\",
            "inputData\\dg.2021-02-01.RData"))

# summary of measured N2O ppm in equilibrated headspace gas
dg %>%
  group_by(sample.source) %>%
  summarize(mean_n2O = mean(n2o.ppm.gc, na.rm = TRUE),
          median_n2o = median(n2o.ppm.gc, na.rm = TRUE)) #mean = 0.36, median = 0.31
```

Simulate a single concentration without measurment error to represent the "true" 
or "known" value for dissolved $\text{N}_2\text{O}$.
```{r dg_calc, echo=TRUE}
# Set volume of syringe/vessel
V <- 140 # 140mL syringe

# Set ratio of gas volume to water volume in the mixing vessel
R_M <- 0.25

B <- 99 # barometric pressure (kPa)
V_g <- V * R_M # volume of reference gas in headspace
V_w <- V * (1 - R_M) # volume of water below headspace
T_c <- 23 # temperature ( C )
M_h <- 0.307 # gas mixture ratio in headspace (Mr_h) * Changes as function of R_M at constant C
M_a_n2o <- 0.310 # gas mixture ratio in reference (Mr_r)
H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
H_ar <- 0.000014 #mol m-3 Pa, range: 0.000013 - 0.000016; Henry's law constant

C <- 1e-6 * B * ( V_g * ( M_h - M_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * M_h )

@@

round( C, 1 ) 
```

```{r dg_calc_vary_r, echo=TRUE}
# Set volume of syringe/vessel
V <- 140 # 140mL syringe

# Set ratio of gas volume to water volume in the mixing vessel
R_M <- 0.25

B <- 99 # barometric pressure (kPa)
V_g <- V * R_M # volume of reference gas in headspace
V_w <- V * (1 - R_M) # volume of water below headspace
T_c <- 23 # temperature ( C )
M_h <- 0.307 # gas mixture ratio in headspace (Mr_h)
M_a_n2o <- 0.310 # gas mixture ratio in reference (Mr_r)
H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
H_ar <- 0.000014 #mol m-3 Pa, range: 0.000013 - 0.000016; Henry's law constant

C <- 1e-6 * B * ( V_g * ( M_h - M_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * M_h )

round( C, 1 ) 
```

## Empricial variation
Estimate variation due to measurement uncertainty using empirical data from
laboratory replicates.
```{r get_gc_error, echo=FALSE}
# READ GC DATA----------------
##########################################################################
#       Code is commented out because collaborators may not have access  #
#       to the file paths. The calculated standard deviation is hard     #
#       coded below.
##########################################################################

@@

Measurement error is defined here as the standard deviation of repeated
analysis of analytical standards throughout the period of time that the
NLA2017 samples were analyzed, which was $\sigma$ = `r n2o_sd_epa`.

## Simulate measurment error

Below, potential errors associated with dissolved $\text{N}_2\text{O}$ 
measurement are simulated and assessed separately for GC observations in:

-

    (1) the equilibrated headspace,

-

    (2) the air, and

-

    (3) the headspace + air

Measurement error is then simulated and assessed separately for:

-

    (4) water temperature,

-

    (5) barometric pressure, and

-

    (5) water volume

Finally, all of the sources (1, 2, 4, 5, and 6) are simulated
simultaneously and assessed as the total variation in dissolved
$\text{N}_2\text{O}$. This total error is then added to the point estimates for 
$\text{N}_2\text{O}$ observed in this study, which results in an estimate of
uncertainty due to measurement error. These error from these simulations was also
carried through to estimated of the saturation ratio. The source-sink status for
each site included in this study was then mapped to present a source-sink 
categorization that reflected these estimated uncertainties. The categorization 
was made such that any observation with simulated error that was within the 
central 95th percentile of

Set the number of simulations for error:

```{r set_n_sims}
nsim <- 1e5
```

### Equilibrated headspace
Simulate measurment error in dissolved $\text{N}_2\text{O}$ due solely to variation in 
measurements within the equlibrated headspace.
```{r simulate_C_h_me, echo=TRUE, message=FALSE}
# Create empty vector to fill below with simulated values for N2O observed with 
# error due to variation in the GC measurments of gas in the equilibrated 
# headspace. The subscript, "me", denotes an observed or measured value, which 
# is contrasted with the known value, C, set above.
C_h_me <- rep(NA, nsim)

# The true value for the concentration of equilibrated gas in the headspace, 
# C_e, was set above. This step simulates additive error in this measurement 
# that is centered over the true value. The observed concentration of
# equilibrated gas is simulated as normally distributed variable, with a mean
# C_e, and a standard deviation specified as the observed sd of replicate GC 
# measurements in our lab. 
M_h_me <- rnorm(nsim, M_h, n2o_sd_epa)

for(i in 1:nsim){
  C_h_me[i] <- 1e-6 * B * (V_g * (M_h_me[i] - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h_me[i])
}

# combine into dataframe for ggplot. 
C_h_df <- tibble(C_h_me = C_h_me, M_h_me = M_h_me)
# save absolute error for plot
C_h_me_abs <- (C_h_me * 1e9) - C
```

The results of the simulations are summarized with density histograms
below.

```{r plot_C_h_me, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_h_df, aes(x = M_h_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = M_h , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurements") +
  xlab("Concentration (nmol L-1)"),

ggplot(C_h_df, aes(x = C_h_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_h_df, aes(x = C_h_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab("Error (nMol)"),
nrow = 1
)
```

### Ambient air
Error in the GC measurements of $\text{N}_2\text{O}$ in the ambient air ($M_a$) is simulated 
below. The result of the simulation is plotted as a histogram to illustrate the
variation in the measurement. THe resulting effect on the uncertainty of the 
dissolved N2O measurement is also plotted along with the absolute error in 
N2O as a result of error in this measurement.
```{r simulate_me_Ca, message=F}
C_a_me <- rep(NA, nsim)

M_a_n2o_me <- rnorm(nsim, M_a_n2o, n2o_sd_epa) # m.e. in measured gas in air

for(i in 1:nsim){
  C_a_me[i] <- 1e-6 * B * (V_g * (M_h - M_a_n2o_me[i]) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_a_df <- tibble(C_a_me = C_a_me, M_a_n2o_me = M_a_n2o_me)
# save absolute error for plot
C_a_me_abs<- (C_a_me * 1e9) - C
```

```{r plot_me_Ca, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_a_df, aes(x = M_a_n2o_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.001 ) +
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = M_a_n2o , color = 'red' ) +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Simulated variation in \nreference gas (Ma) GC \nmeasurements") +
  xlab( "Concentration (nmol L-1)"),

ggplot(C_a_df, aes(x = C_a_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_a_df, aes(x = C_a_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1)
```

### Water temperature

```{r simulate_me_T_c, message=F}
C_tc_me <- rep(NA, nsim)

# How to get standard deviation of temperature measurement?
# calibration certificate reports an uncertainty of +/-0.058C
# take standard deviation of simulated measurements with specified uncertainty
T_c_sd <- sd(runif(n = nsim, min = T_c - 0.058, max = T_c + 0.058)) # calibration certificate reports an uncertainty of +/-0.058C
T_c_me <- rnorm(nsim, T_c, T_c_sd) #m.e. in measured water temp

for(i in 1:nsim){
  C_tc_me[i] <- 1e-6 * B * (V_g * (M_h - M_a_n2o) / (8.3144598 * (T_c_me[i] + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)) * M_h)
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_tc_df <- tibble(C_tc_me = C_tc_me, T_c_me = T_c_me)
# save absolute error for plot
C_tc_me_abs <- (C_tc_me * 1e9) - C
```

```{r plot_me_T_c, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_tc_df, aes(x = T_c_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = T_c , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in water \ntemperature measurements \n") +
  xlab("Temperature (C)"),

ggplot(C_tc_df, aes(x = C_tc_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_tc_df, aes(x = C_tc_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab("Error (nMol)"),
nrow = 1)
```

### Barometric pressure

```{r simulate_me_B, message=F}
C_b_me <- rep(NA, nsim)

# How to get standard deviation of barometric pressure?
# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa), but EPA calibration
# checks performed in the AWBERC metrology lab demonstrate much greater precision.
# below we calculat sd of replicate measurements made with 15 different barometers.
# we then take the mean of the 15 sd values.

#2023 barometer verification in AWBERC metrology lab
B_sd <- mean(
  sd(c(999,998.6)*0.1), # YSI ProDSS
  sd(c(1000.848, 1000.848)*0.1), # EXO 21G104022
  sd(c(1000.982, 1000.982)*0.1), # EXO 18H111572
  sd(c(1000.982, 1000.982)*0.1), # EXO 22F105975
  sd(c(1001.382, 1001.382)*0.1), # EXO 21K100162
  sd(c(1000.98, 1000.85)*0.1), # EXO 21F103579
  sd(c(1000.31, 1000.44)*0.1), # MDS 16D100327
  sd(c(1001.25, 1001.25)*0.1), # MDS 12F102618
  sd(c(1000, 1000)*0.1), # MDS 16D100328
  sd(c(1002, 1002)*0.1), # MDS 02K599 AF
  sd(c(1003,1003)*0.1), # MDS 12J0793 AF
  sd(c(1000.85,1000.85)*0.1), # MDS 16E100831
  sd(c(1000.32, 1000.32)*0.1), # MDS 12F102619
  sd(c(1002, 1002)*0.1)) # MDS 15F104725

B_me <- rnorm(nsim, B, B_sd) #m.e. in measured barometric pressure

for(i in 1:nsim){
  C_b_me[i] <- 1e-6 * B_me[i] * (V_g * (M_h - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_b_df <- tibble(C_b_me = C_b_me, B_me = B_me)
# save absolute error for plot
C_b_me_abs <- (C_b_me * 1e9) - C
```

```{r plot_me_B, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_b_df, aes(x = B_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = B , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in \nbarometric pressure \nmeasurements") +
  xlab("Barometric pressure (kPa)"),
